---
title: Boxplot with trend line
author: JW
date: '2018-04-13'
slug: boxplot-with-trend-line
categories:
  - R
  - ggplot
tags:
  - r
  - ggplot
  - plot
---

## Background

Last week my colleague showed me a drawn graph and asked can we plot this kind of graph. It is a graph with grouped boxplot. Additionally, it contains a trend line that connects mean of each data box. So this post explains how to draw a graph like that using two ways;

+ **r** *ggplot2* library 
+ **r** *base plots* (used codes from The R Graph Gallery - https://www.r-graph-gallery.com/146-boxplot-for-time-series/)

## The data

This data is sample data set from my colleague. The data contains NDVI (Normalize Different Vegetation Index) time series values for three different crops (Maize, Potatoes, Sugar beet) extracted from Landsat 8 data.

### Load data

#### Load libraries

```{r echo = TRUE, error = FALSE, warning = FALSE, message = FALSE}
library(ggplot2)
library(readr)
library(plyr)
library(dplyr)
library(RColorBrewer)
```

#### Read and prepare data
```{r echo = TRUE, error = FALSE, warning = FALSE, message = FALSE}
read.df <- read.csv("D:/Blog_data/later_summercrop_ts.csv", header = TRUE)
sample.df <- read.df # Make a copy of the read table

levels(sample.df$date) # Checking the levels in the date column
levels(sample.df$crop) # Checking the levels in the crop column

# Reorder levels
sample.df$date <- factor(sample.df$date, levels = c("15-Apr-15", "24-Apr-15", "10-May-15", "04-Jul-15","07-Aug-15", "21-Aug-15", "23-Aug-15", "30-Aug-15"))

sample.df$crop <- factor(sample.df$crop, levels = c("maize", "potato", "sugar beet"))
```

### Plot with *ggplot* library

```{r echo = TRUE, error = FALSE, warning = FALSE, message = FALSE, fig.width=10, fig.height=4}

# Create a color palette for three crops
my_colors = c("greenyellow", "gold3", "darksalmon")

# Create a plot
ggplot(data=sample.df, aes(x=date, y=NDVI, fill=crop)) +
  geom_boxplot() + ylim(c(-0.2,1.2)) +
  stat_summary(fun.y=mean, colour="red",
               geom="point",position=position_dodge(width=0.75)) +
  stat_summary(fun.y=mean, colour="red", aes(group=1),
               geom="line", lwd=1, lty=1) +
  theme_bw(base_size = 12) +
  scale_fill_manual(values = my_colors, name = "Crop Type", labels = c("Maize","Potato","Sugar beet")) +
  labs(x="Date", y="NDVI", 
       title="Phenology of the late summer crops", 
       subtitle="NDVI extracted from Landsat 8 time series data")
```

### Plot with *base* library

```{r echo = TRUE, error = FALSE, warning = FALSE, message = FALSE, fig.width=10, fig.height=4}
# Make the boxplot
boxplot(sample.df$NDVI ~ sample.df$crop+sample.df$date , ylim=c(-0.2,1.2), xaxt="n" , xlab="Date" , col=my_colors , pch=20 , cex=0.3 , 
        ylab="NDVI", xlab="Date", main="Phenology of the late summer crops \n NDVI extracted from Landsat 8 time series data" )
abline(v= seq(0, 3*8, 3) + 0.5 , col="grey")
axis(1, labels=levels(sample.df$date) , at=seq(2,3*8,3) )


# Add general trend
a=aggregate(sample.df$NDVI , by=list(sample.df$crop, sample.df$date) , mean)
lines(a[,3], type="l" , col="red" , lwd=2)


#Add legend for crops
legend("top", fill = my_colors, legend = c("Maize","Potato","Sugar beet"), horiz = T)
```
